---
title: "Modeling"
author: "Gene Estrada"
date: "2024-02-14"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)

setwd("~/leech_indicators")

library(tidyverse)
library(glmmTMB)
library(ggmap)
library(sf)
library(ggstatsplot)
library(sjPlot)
library(car)
library(PerformanceAnalytics)
library(performance)
library(gridExtra)
library(MuMIn)
library(cowplot)
library(Hmisc)

#read in pruned mammal observation table
pruned <- readRDS("data/pruned.rds")

#ordering forest types and forest type partitions
pruned$habitat <- factor(pruned$habitat, levels = c("Peat Swamp", "Freshwater Swamp", "Alluvial Bench", 
                                          "Lowland Sandstone", "Lowland Granite", "Upland Granite", "Montane"))

pruned$partition <- factor(pruned$partition, levels = c("PS1", "FS1", "AB1", "AB2", "LS1", "LS2", 
                                                        "LG1", "LG2", "UG1", "UG2", "MO1", "MO2"))

#create modified partition column, combining UG and MO partitions
pruned <- pruned %>%
  mutate(partition.mod = case_when(partition == "UG1" ~ "UGM1",
                                   partition == "MO1" ~ "UGM1",
                                   partition == "UG2" ~ "UGM2",
                                   partition == "MO2" ~ "UGM2",
                                   TRUE ~ partition))

pruned$partition.mod <- factor(pruned$partition.mod, levels = c("PS1", "FS1", "AB1", "AB2", "LS1", "LS2", 
                                                        "LG1", "LG2", "UGM1", "UGM2"))

#add pheno data by modified partition
pheno_by_pt_mod <- read.csv(file = "data/pheno.by.ft.raw.csv", header = TRUE)

#changing naming convention
pheno_by_pt_mod <- pheno_by_pt_mod %>%
  mutate(partition.mod = case_when(partishun == "PS.I" ~ "PS1",
                                   partishun == "FS.I" ~ "FS1",
                                   partishun == "AB.I" ~ "AB1",
                                   partishun == "AB.II" ~ "AB2",
                                   partishun == "LS.I" ~ "LS1",
                                   partishun == "LS.II" ~ "LS2",
                                   partishun == "LG.I" ~ "LG1",
                                   partishun == "LG.II" ~ "LG2",
                                   partishun == "UGM.I" ~ "UGM1",
                                   partishun == "UGM.II" ~ "UGM2",
                                   TRUE ~ partishun))

#add new pheno data to pruned obs table
pruned <- left_join(pruned, pheno_by_pt_mod[,c("partition.mod", "z_pheno_mod", "samplingperiod")], 
          by = c("partition.mod", "samplingperiod"))

#read in mammal observation by location summary table
mammals_by_loc <- read.csv(file = "data/mammals_by_loc.csv", header = TRUE)

#read in table describing camera trap nights by ct location, adding to mammal summary table
act_days <- read.csv(file = "data/act_days_by_location.csv", header = TRUE)
mammals_by_loc <- left_join(mammals_by_loc, act_days[,-1], by = "locationID")

#read in mammal trait data
mammal_traits <- read.csv(file = "data/mammal_list_ct_trait_202403.csv", header = TRUE)

pruned <- left_join(pruned, 
                    mammal_traits[,c("Species", "ActivityCycle", "AdultBodyMass_g", 
                                     "Terrestriality", "TrophicLevel", "Diet")],
                  by = "Species")

#aggregating observations by sampling period
pruned_aggregated <- pruned %>%
  group_by(samplingperiod, Species, locationID) %>%
  summarise(Number.of.Animals = sum(num_observed), .groups = "drop")

pruned_aggregated <- pruned %>%
  distinct(samplingperiod, locationID, .keep_all = TRUE) %>%
  select(location, locationID, samplingperiod, habitat, partition.mod, on_off_trail, 
         altitude, z_rain, z_min, z_max, z_pheno_mod, z_elev, z_picta_part, 
         z_zeylanica_part, active.days) %>%
  left_join(pruned_aggregated, ., by = c("samplingperiod", "locationID"))

pruned_aggregated <- left_join(pruned_aggregated, 
                    mammal_traits[,c("Species", "ActivityCycle", "AdultBodyMass_g", "Family",
                                     "Terrestriality", "TrophicLevel", "Diet")],
                  by = "Species")

#removing observations from kerangas and rangkong forest types
pruned_aggregated <- pruned_aggregated[!is.na(pruned_aggregated$habitat),]

#removing arboreal mammal observations (keeping unid rats and M.nemestrina)
pruned_aggregated$Terrestriality[pruned_aggregated$Species == "Unid Rat"] <- 1
pruned_aggregated$Terrestriality[pruned_aggregated$Species == "Macaca nemestrina"] <- 1

pruned_aggregated <- pruned_aggregated %>%
  filter(Terrestriality == 1)

pruned_aggregated$on.off.trail <- pruned_aggregated$on_off_trail
```


Creating mammal observation table by family
```{r}
mammals_by_fam <- pruned %>%
  group_by(Family, Species) %>%
  summarise(count = sum(num_observed)) %>%
  arrange(Family, desc(count)) %>%
  print(n = Inf)
```


Spearman correlation between quantitative predictors
```{r eval=FALSE, include=FALSE}
chart.Correlation(pruned_aggregated[,c("z_rain","z_max","z_min","z_pheno_mod","z_elev",
                            "z_picta_part", "z_zeylanica_part")],
                  histogram = TRUE, pch = 16, method = "spearman")
```


## Mammal traits

Active period
1 = nocturnal, 2 = nocturnal or diurnal/crepuscular, 3 = diurnal
```{r}
table(mammal_traits$ActivityCycle)

#missing data for observations without species ID (e.g., Unid squirrel)
table(is.na(mammal_traits$ActivityCycle))
mammal_traits[is.na(mammal_traits$ActivityCycle),]
```

Body mass
```{r}
hist(mammal_traits$AdultBodyMass_g, breaks = 30)
hist(mammal_traits$AdultBodyMass_g[mammal_traits$AdultBodyMass_g < 20000], breaks = 30)
summary(mammal_traits$AdultBodyMass_g)

#some missing data here - mostly Unid species or animals not identified to species level
table(is.na(mammal_traits$AdultBodyMass_g))
mammal_traits[is.na(mammal_traits$AdultBodyMass_g),]
```

Terrestriality
1 = terrestrial, 2 = arboreal
```{r}
table(mammal_traits$Terrestriality)

#missing values
table(is.na(mammal_traits$Terrestriality))
mammal_traits[is.na(mammal_traits$AdultBodyMass_g),]
```

Trophic Level
1 = herbivore, 2 = omnivore, 3 = carnivore
```{r}
table(mammal_traits$TrophicLevel)

#missing values
table(is.na(mammal_traits$TrophicLevel))
mammal_traits[is.na(mammal_traits$TrophicLevel),]
```

## Modeling all mammals

Correlation between predictors in full model
```{r}
df <- pruned_aggregated[, c("z_picta_part", "z_zeylanica_part", 
                            "z_rain", "z_min", "z_max", "z_pheno_mod")]

colnames(df) <- c("tiger leech", "brown leech", "rainfall", 
                  "min. temperature", "max. temperature", "fruit availability")

#correlation plot
corr_results <- rcorr(as.matrix(df), type = "spearman")

cor_matrix <- corr_results$r
p_matrix <- corr_results$P 

#save plot
#jpeg("SpearCorrPlot.jpg", width = 5, height = 6, units = "in", res = 350)

#corrplot(cor_matrix,
 #        method = "color",          
  #       type = "upper",            
   #      tl.col = "black",          
    #     tl.srt = 45,               
     #    tl.cex = 0.8,              
      #   cl.pos = "r",              
       #  addCoef.col = TRUE,
        # number.cex = 1,
         #number.font = 1,
         #diag = FALSE)

#dev.off()
```


Zero-inflated model using only leech counts, by partition
```{r}
#model
mod1.1 <- glmmTMB(Number.of.Animals ~
         z_picta_part +              #rate of H. picta by partition
         z_zeylanica_part +          #rate of H. zeylanica by partition
         samplingperiod +            #30-day sampling period ID
         on.off.trail +              #CT placement on- or off-trail
         (1|locationID) +            #controlling for CT locations
         (1|Species),                #controlling for species 
         offset = log(active.days),  #controlling for uneven CT survey effort 
         data = pruned_aggregated,
         family = nbinom2,           #negative binomial
         ziformula = ~1)             #intercept-only zero-inflated portion

#summary
summary(mod1.1)

#check Variance Inflation Factor values (VIF>5 indicates multicollinearity issues)
check_collinearity(mod1.1)

#coefficient plot
ggcoefstats(mod1.1, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  scale_y_discrete(labels = c("tiger leeches", "brown leeches", "sampling period", "CT on/off trail")) +
  labs(title = "All Mammals - Leech Counts by partition")

#marginal effects plots for leech species predictors
mp1 <- plot_model(mod1.1, type = "pred", terms = "z_picta_part") 
mp2 <- plot_model(mod1.1, type = "pred", terms = "z_zeylanica_part")
grid.arrange(mp1, mp2, nrow = 1)
```

Model with leeches and ecological predictors
```{r}
#model using {glmmTMB}
mod2 <- glmmTMB(Number.of.Animals ~
         z_picta_part +
         z_zeylanica_part +
         samplingperiod +
         on.off.trail +
         z_rain +             #rainfall     
         z_max +              #max temp     
         z_min +              #min temp     
         z_pheno_mod +        #ripe fruit abundance     
         (1|locationID) +
         (1|Species),
         offset = log(active.days),
         data = pruned_aggregated,
         family = nbinom2,
         ziformula = ~1)

summary(mod2)
check_collinearity(mod2)

#plot
broom.mixed::tidy(mod2, conf.int = TRUE) %>%
  filter(effect == "fixed", !term %in% c("samplingperiod", "on.off.trailON", "(Intercept)")) %>%
  mutate(term = fct_reorder(term, estimate, .desc = FALSE)) %>%
  ggplot(aes(x = term, y = estimate)) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0, color = "gray60") +
  geom_errorbar(aes(ymin = conf.low + (conf.high - conf.low) * 0.25, 
                    ymax = conf.high - (conf.high - conf.low) * 0.25), 
                width = 0, color = "black", size = 1) +
  geom_point(size = 2, color = "black") +
  coord_flip() +  
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  scale_x_discrete(labels = c("z_picta_part" = "tiger leech",
                              "z_zeylanica_part" = "brown leech",
                              "z_rain" = "rainfall",
                              "z_max" = "max. temp",
                              "z_min" = "min. temp",
                              "z_pheno_mod" = "fruit")) +
  theme_classic() +
  labs(y = "estimate") +
  theme(axis.title.y = element_blank())


#marginal effects plots for leech species predictors
mp1 <- plot_model(mod2, type = "pred", terms = "z_picta_part") +
  coord_cartesian(ylim = c(0, 0.016)) +
  theme_classic() +
  labs(x = "tiger leech rate (z)", y = "mammal count",  title = "")
mp2 <- plot_model(mod2, type = "pred", terms = "z_zeylanica_part") +
  coord_cartesian(ylim = c(0, 0.016)) +
  theme_classic() +
  theme(axis.title.y = element_blank(), axis.text.y = element_blank()) +
  labs(x = "brown leech rate (z)", y = "", title = "")

plot_grid(mp1 + theme(legend.position = "none"), mp2, 
          labels = c("A", "B"),
          label_size = 14, label_fontface = "bold",
          rel_widths = c(0.52, 0.48))

#ggsave("AllMammals_MEPlot.jpg", width = 6.5, height = 3.5, dpi = 350)
```

Model with ecological predictors only
```{r}
mod3 <- glmmTMB(Number.of.Animals ~
         samplingperiod +
         on.off.trail +
         z_rain +                  
         z_max +                   
         z_min +      
         z_pheno_mod +   
         (1|locationID) +
         (1|Species),
         offset = log(active.days),
         data = pruned_aggregated,
         family = nbinom2,
         ziformula = ~1)

summary(mod3)

ggcoefstats(mod3, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  scale_y_discrete(labels = c("sampling period", "on/off trail", "rainfall",
                              "max temp", "min temp", "fruit")) +
  labs(title = "All Mammals - All Predictors Except Leeches")
```

Full model with interactions between leeches and rain
```{r}
mod4 <- glmmTMB(Number.of.Animals ~
         z_picta_part * z_rain +
         z_zeylanica_part * z_rain +
         samplingperiod +
         on.off.trail +
         z_max +                   
         z_min +                   
         z_pheno_mod +             
         (1|locationID) +
         (1|Species),
         offset = log(active.days),
         data = pruned_aggregated,
         family = nbinom2,
         ziformula = ~1)

summary(mod4)

#coef plot
broom.mixed::tidy(mod4, conf.int = TRUE) %>%
  filter(effect == "fixed", !term %in% c("samplingperiod", "on.off.trailON", "(Intercept)")) %>%
  mutate(term = fct_reorder(term, estimate, .desc = FALSE)) %>%
  ggplot(aes(x = term, y = estimate)) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0, color = "gray20") +
  geom_errorbar(aes(ymin = conf.low + (conf.high - conf.low) * 0.25, 
                    ymax = conf.high - (conf.high - conf.low) * 0.25), 
                width = 0, color = "black", size = 1.1) +
  geom_point(size = 2, color = "black") +
  coord_flip() +  
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey60") +
  scale_x_discrete(labels = c("z_picta_part" = "tiger leech",
                              "z_zeylanica_part" = "brown leech",
                              "z_rain" = "rainfall",
                              "z_max" = "max. temperature",
                              "z_min" = "min. temperature",
                              "z_pheno_mod" = "fruit abundance",
                              "z_picta_part:z_rain" = "tiger leech X rain",
                              "z_rain:z_zeylanica_part" = "brown leech X rain")) +
  theme_classic() +
  labs(y = "coefficient estimate", title = "all-mammals model") +
  theme(axis.title.y = element_blank())

#ggsave(filename = "AllMammalsInt_coefplot.jpg", device = "jpeg", width = 5, height = 3.5, dpi = 300)
```

interaction plot
```{r}
p1 <- plot_model(mod4, type = "pred", terms = c("z_picta_part", "z_rain [minmax]")) +
  theme_classic() +
  labs(x = "tiger leech rate (z)", y = "mammal count", 
       title = "", color = "rainfall (z)") +
  theme(legend.position = "none") +
  coord_cartesian(ylim = c(0, 0.02))

p2 <- plot_model(mod4, type = "pred", terms = c("z_zeylanica_part", "z_rain [minmax]")) +
  theme_classic() +
  labs(x = "brown leech rate (z)", y = "", 
       title = "", color = "rainfall (z)") +
  theme(legend.position = "right", axis.text.y  = element_blank()) +
  coord_cartesian(ylim = c(0, 0.02))

plot_grid(p1 + theme(legend.position = "none"), p2, 
          labels = c("A", "B"),
          label_size = 14, label_fontface = "bold",
          rel_widths = c(0.45, 0.55))

#ggsave("AllMammals_IntPlot.jpg", width = 7.5, height = 3.5, dpi = 350)
```

model selection with AIC
```{r}
AIC(mod1.1, mod2, mod3, mod4)
```


## Modeling single mammal species

Table showing number of observations by species over study period.
```{r}
mammals_by_obs <- as.data.frame(table(pruned$Species)) %>% arrange(desc(Freq)) %>% print()
```


### Rats (Family Muridae)

```{r}
hist(pruned$num_observed[pruned$Species == "Unid Rat"])
table(pruned$num_observed[pruned$Species == "Unid Rat"]) # >99% of obs are solitary rats
```

model - leech counts only 
```{r}
rats_m1 <- glmmTMB(Number.of.Animals ~
         z_picta_part +
         z_zeylanica_part +
         samplingperiod +   
         on.off.trail +   
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$Species == "Unid Rat",],
         family = nbinom2,
         ziformula = ~1)

summary(rats_m1)

ggcoefstats(rats_m1, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  scale_y_discrete(labels = c("tiger leeches", "brown leeches", "sampling period", "on/off trail")) +
  labs(title = "Family Muridae (rats) - Leech Counts by Partition")
```

model - leech counts plus ecological predictors
```{r}
rats_m2 <- glmmTMB(Number.of.Animals ~
         z_picta_part +
         z_zeylanica_part +
         samplingperiod +
         z_rain +                  
         z_max +                   
         z_min +
         z_pheno_mod +
         on.off.trail +
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$Species == "Unid Rat",],
         family = nbinom2,
         ziformula = ~1)

summary(rats_m2)
check_collinearity(rats_m2)

ggcoefstats(rats_m2, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  scale_y_discrete(labels = c("tiger leeches", "brown leeches", "sampling period", "rainfall",
                              "max temp", "min temp", "fruit", "on/off trail")) +
  labs(title = "Family Muridae (rats) - All Predictors")
```

```{r}
rats_m3 <- glmmTMB(Number.of.Animals ~
         z_picta_part * z_rain +
         z_zeylanica_part * z_rain +
         samplingperiod +
         z_rain +                  
         z_max +                   
         z_min +
         z_pheno_mod +
         on.off.trail +
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$Species == "Unid Rat",],
         family = nbinom2,
         ziformula = ~1)

summary(rats_m3)
check_collinearity(rats_m3)

ggcoefstats(rats_m3, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  labs(title = "Family Muridae (rats) - All Predictors")
```

Ecological predictors only
```{r}
rats_m4 <- glmmTMB(Number.of.Animals ~
         samplingperiod +
         z_rain +                  
         z_max +                   
         z_min +
         z_pheno_mod +
         on.off.trail +
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$Species == "Unid Rat",],
         family = nbinom2,
         ziformula = ~1)
```

compare models
```{r}
#comparing models using AIC
AIC(rats_m1, rats_m2, rats_m3, rats_m4)
```


### Bearded pigs - Sus barbatus

```{r}
hist(pruned$num_observed[pruned$Species == "Sus barbatus"])
table(pruned$num_observed[pruned$Species == "Sus barbatus"])
```

model - leech counts only 
```{r}
pigs_m1 <- glmmTMB(Number.of.Animals ~
         z_picta_part +
         z_zeylanica_part +
         samplingperiod + 
         on.off.trail +
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$Species == "Sus barbatus",],
         family = nbinom2,
         ziformula = ~1)

summary(pigs_m1)

ggcoefstats(pigs_m1, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  scale_y_discrete(labels = c("tiger leeches", "brown leeches", "sampling period", "on/off trail")) +
  xlim(-1.5, 0.5) + 
  labs(title = "Sus barbatus - Leech Predictors Only")
```

model - leech counts plus ecological predictors
```{r}
pigs_m2 <- glmmTMB(Number.of.Animals ~
         z_picta_part +
         z_zeylanica_part +
         samplingperiod +
         z_rain +                  
         z_max +                   
         z_min +                   
         z_pheno_mod +
         on.off.trail +
         (1|partition.mod) +
         offset(log(active.days)),
         data = pruned_aggregated[pruned_aggregated$Species == "Sus barbatus",],
         family = nbinom2,
         ziformula = ~1)

summary(pigs_m2)
check_collinearity(pigs_m2)

ggcoefstats(pigs_m2, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  scale_y_discrete(labels = c("tiger leeches", "brown leeches", "sampling period", "rainfall",
                              "max temp", "min temp", "fruit", "on/off trail")) +
  labs(title = "Sus barbatus - All Predictors")
```

full model with interactions
```{r}
pigs_m3 <- glmmTMB(Number.of.Animals ~
         z_picta_part * z_rain +
         z_zeylanica_part * z_rain +
         samplingperiod +
         z_max +                   
         z_min +                   
         z_pheno_mod +
         on.off.trail +
         (1|partition.mod) +
         offset(log(active.days)),
         data = pruned_aggregated[pruned_aggregated$Species == "Sus barbatus",],
         family = nbinom2,
         ziformula = ~1)

summary(pigs_m3)
check_collinearity(pigs_m3)

ggcoefstats(pigs_m3, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  labs(title = "Sus barbatus - All Predictors")
```

ecological only
```{r}
pigs_m4 <- glmmTMB(Number.of.Animals ~
         samplingperiod +
         z_rain +                  
         z_max +                   
         z_min +                   
         z_pheno_mod +
         on.off.trail +
         (1|partition.mod) +
         offset(log(active.days)),
         data = pruned_aggregated[pruned_aggregated$Species == "Sus barbatus",],
         family = nbinom2,
         ziformula = ~1)
```

compare models
```{r}
AIC(pigs_m1, pigs_m2, pigs_m3, pigs_m4)
```

### Malayan Civet - Viverra tangalunga

var < mean, underdispersed
```{r}
hist(pruned$num_observed[pruned$Species == "Viverra tangalunga"])
table(pruned$num_observed[pruned$Species == "Viverra tangalunga"])
```

model - leech counts only 
```{r}
vtan_m1 <- glmmTMB(Number.of.Animals ~
         z_picta_part +
         z_zeylanica_part +
         samplingperiod +
         on.off.trail +
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$Species == "Viverra tangalunga",],
         family = nbinom2,
         ziformula = ~1)

summary(vtan_m1)

ggcoefstats(vtan_m1, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  scale_y_discrete(labels = c("tiger leeches", "brown leeches", "sampling period", "on/off trail")) +
  labs(title = "Viverra tangalunga - Leech Predictors Only")
```

model - leech counts plus ecological predictors
```{r}
vtan_m2 <- glmmTMB(Number.of.Animals ~
         z_picta_part +
         z_zeylanica_part +
         samplingperiod +
         z_rain +                  
         z_max +                   
         z_min + 
         z_pheno_mod +
         on.off.trail +
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$Species == "Viverra tangalunga",],
         family = nbinom2,
         ziformula = ~1)

summary(vtan_m2)
check_collinearity(vtan_m2)

ggcoefstats(vtan_m2, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  scale_y_discrete(labels = c("tiger leeches", "brown leeches", "sampling period", "rainfall",
                              "max temp", "min temp", "fruit", "on/off trail")) +
  labs(title = "Viverra tangalunga - All Predictors")
```

full model with interactions
```{r}
vtan_m3 <- glmmTMB(Number.of.Animals ~
         z_picta_part * z_rain +
         z_zeylanica_part * z_rain +
         samplingperiod +
         z_max +                   
         z_min + 
         z_pheno_mod +
         on.off.trail +
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$Species == "Viverra tangalunga",],
         family = nbinom2,
         ziformula = ~1)

summary(vtan_m3)
check_collinearity(vtan_m3)

ggcoefstats(vtan_m3, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  labs(title = "Viverra tangalunga - All Predictors")
```

```{r}
vtan_m4 <- glmmTMB(Number.of.Animals ~
         samplingperiod +
         z_rain +                  
         z_max +                   
         z_min + 
         z_pheno_mod +
         on.off.trail +
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$Species == "Viverra tangalunga",],
         family = nbinom2,
         ziformula = ~1)
```


compare models
```{r}
AIC(vtan_m1, vtan_m2, vtan_m3, vtan_m4)
```


### Java Mouse Deer - Tragulus javanicus

```{r}
hist(pruned$num_observed[pruned$Species == "Tragulus javanicus"])
table(pruned$num_observed[pruned$Species == "Tragulus javanicus"])
```

model - leech counts only. No reliable effect for either leech species
```{r}
tjav_m1 <- glmmTMB(Number.of.Animals ~
         z_picta_part +
         z_zeylanica_part +
         samplingperiod +
         on.off.trail +
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$Species == "Tragulus javanicus",],
         family = nbinom2,
         ziformula = ~1)

summary(tjav_m1)

ggcoefstats(tjav_m1, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  scale_y_discrete(labels = c("tiger leeches", "brown leeches", "sampling period", "on/off trail")) +
  labs(title = "Tragulus javanicus - Leech Predictors Only")
```

model - leech counts plus ecological predictors. No reliable effects
```{r}
tjav_m2 <- glmmTMB(Number.of.Animals ~
         z_picta_part +
         z_zeylanica_part +
         samplingperiod +
         z_rain +                  
         z_max +                   
         z_min +                   
         z_pheno_mod +
         on.off.trail +
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$Species == "Tragulus javanicus",],
         family = nbinom2,
         ziformula = ~1)

summary(tjav_m2)
check_collinearity(tjav_m2)

ggcoefstats(tjav_m2, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  scale_y_discrete(labels = c("tiger leeches", "sampling period", "rainfall",
                              "max temp", "min temp", "fruit", "on/off trail")) +
  labs(title = "Tragulus javanicus - All Predictors")
```

model with interactions
```{r}
tjav_m3 <- glmmTMB(Number.of.Animals ~
         z_picta_part * z_rain +
         #z_zeylanica_part * z_rain +  #highly correlated (VIF > 14)
         samplingperiod +
         z_max +                   
         z_min +                   
         z_pheno_mod +
         on.off.trail +
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$Species == "Tragulus javanicus",],
         family = nbinom2,
         ziformula = ~1)

summary(tjav_m3)
check_collinearity(tjav_m3)

ggcoefstats(tjav_m3, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  labs(title = "Tragulus javanicus - All Predictors")
```

ecological predictors only
```{r}
tjav_m4 <- glmmTMB(Number.of.Animals ~
         samplingperiod +
         z_rain +                  
         z_max +                   
         z_min +                   
         z_pheno_mod +
         on.off.trail +
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$Species == "Tragulus javanicus",],
         family = nbinom2,
         ziformula = ~1)
```

compare models
```{r}
AIC(tjav_m1, tjav_m2, tjav_m3, tjav_m4)
```

interaction plot
```{r}
plot_model(tjav_m3, type = "pred", terms = c("z_picta_part", "z_rain [minmax]")) +
  theme_classic() +
  labs(x = "tiger leech rate (z)", y = "mammal count", 
       title = "leech X rainfall interaction", color = "rainfall (z)")
```


### Pig-tailed macaque - Macaca nemestrina

```{r}
hist(pruned$num_observed[pruned$Species == "Macaca nemestrina"])
table(pruned$num_observed[pruned$Species == "Macaca nemestrina"])
```

model - leech counts only. No reliable effect of either leech species
```{r}
mnem_m1 <- glmmTMB(Number.of.Animals ~
         z_picta_part +
         z_zeylanica_part +
         samplingperiod +
         on.off.trail +
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$Species == "Macaca nemestrina",],
         family = nbinom2,
         ziformula = ~1)

summary(mnem_m1)

ggcoefstats(mnem_m1, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  scale_y_discrete(labels = c("tiger leeches", "brown leeches", "sampling period", "on/off trail")) +
  labs(title = "Macaca nemestrina - Leech Predictors Only")
```

model - leech counts plus ecological predictors
```{r}
mnem_m2 <- glmmTMB(Number.of.Animals ~
         z_picta_part +
         z_zeylanica_part +
         samplingperiod +
         z_rain +                  
         z_max +                   
         z_min +                   
         z_pheno_mod + 
         on.off.trail +
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$Species == "Macaca nemestrina",],
         family = nbinom2,
         ziformula = ~1)

summary(mnem_m2)
check_collinearity(mnem_m2)

ggcoefstats(mnem_m2, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  scale_y_discrete(labels = c("tiger leeches", "brown leeches", "sampling period", "rainfall",
                              "max temp", "min temp", "fruit", "on/off trail")) +
  labs(title = "Macaca nemestrina - All Predictors")
```

model with interactions
```{r}
mnem_m3 <- glmmTMB(Number.of.Animals ~
         z_picta_part * z_rain +
         z_zeylanica_part * z_rain +
         samplingperiod +
         z_max +                   
         z_min +                   
         z_pheno_mod + 
         on.off.trail +
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$Species == "Macaca nemestrina",],
         family = nbinom2,
         ziformula = ~1)

summary(mnem_m3)
check_collinearity(mnem_m3)

ggcoefstats(mnem_m3, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  labs(title = "Macaca nemestrina - All Predictors")
```

ecological only
```{r}
mnem_m4 <- glmmTMB(Number.of.Animals ~
         samplingperiod +
         z_rain +                  
         z_max +                   
         z_min +                   
         z_pheno_mod + 
         on.off.trail +
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$Species == "Macaca nemestrina",],
         family = nbinom2,
         ziformula = ~1)
```

compare models
```{r}
AIC(mnem_m1, mnem_m2, mnem_m3, mnem_m4)
```


## Modeling by groups

### Grouping by family

#### Cervidae

Two species of Muntiacus, 1,372 total individuals, mostly M.muntjak (~60%)
```{r}
#n observations by species and by family
table(pruned$Species[pruned$Family == "Cervidae" & pruned$num_observed != 0])
length(pruned$Species[pruned$Family == "Cervidae" & pruned$num_observed != 0])

table(pruned$habitat[pruned$Family == "Cervidae" & pruned$num_observed != 0])
table(pruned$partition[pruned$Family == "Cervidae" & pruned$num_observed != 0])

hist(pruned$num_observed[pruned$Family == "Cervidae"])
table(pruned$num_observed[pruned$Family == "Cervidae"])
```

model - leech predictors
```{r}
cerv_m1 <- glmmTMB(Number.of.Animals ~
         z_picta_part +
         z_zeylanica_part +
         samplingperiod +
         on.off.trail +
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$Family == "Cervidae",],
         family = nbinom2,
         ziformula = ~1)

summary(cerv_m1)

ggcoefstats(cerv_m1, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  scale_y_discrete(labels = c("tiger leeches", "brown leeches", "sampling period", "on/off trail")) +
  labs(title = "Cervidae - Leech Predictors Only")
```

model - leech counts plus ecological predictors
```{r}
cerv_m2 <- glmmTMB(Number.of.Animals ~
         z_picta_part +
         z_zeylanica_part +
         samplingperiod +
         on.off.trail +
         z_rain +                  
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod) +
         offset(log(active.days)),
         data = pruned_aggregated[pruned_aggregated$Family == "Cervidae",],
         family = nbinom2,
         ziformula = ~1)

summary(cerv_m2)
check_collinearity(cerv_m2)

ggcoefstats(cerv_m2, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  scale_y_discrete(labels = c("tiger leeches", "brown leeches", "sampling period", "on/off trail",
                              "rainfall", "max temp", "min temp", "fruit")) +
  labs(title = "Cervidae - All Predictors")
```

model - full plus interactions
```{r}
cerv_m3 <- glmmTMB(Number.of.Animals ~
         z_picta_part * z_rain +
         z_zeylanica_part * z_rain +
         samplingperiod +
         on.off.trail +
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod) +
         offset(log(active.days)),
         data = pruned_aggregated[pruned_aggregated$Family == "Cervidae",],
         family = nbinom2,
         ziformula = ~1)

summary(cerv_m3)
check_collinearity(cerv_m3)

ggcoefstats(cerv_m3, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  labs(title = "Cervidae - All Predictors + interactions")
```

ecological only
```{r}
cerv_m4 <- glmmTMB(Number.of.Animals ~
         samplingperiod +
         on.off.trail +
         z_rain +                  
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod) +
         offset(log(active.days)),
         data = pruned_aggregated[pruned_aggregated$Family == "Cervidae",],
         family = nbinom2,
         ziformula = ~1)
```

model comparison
```{r}
AIC(cerv_m1, cerv_m2, cerv_m3, cerv_m4)
```


#### Felidae

Five species of Felids, 506 total individuals, mostly Neofelis (71%)
```{r}
#n observations by species and by family
table(pruned$Species[pruned$Family == "Felidae" & pruned$num_observed != 0])
length(pruned$Species[pruned$Family == "Felidae" & pruned$num_observed != 0])

table(pruned$habitat[pruned$Family == "Felidae" & pruned$num_observed != 0])
table(pruned$partition[pruned$Family == "Felidae" & pruned$num_observed != 0])

hist(pruned$num_observed[pruned$Family == "Felidae"])
table(pruned$num_observed[pruned$Family == "Felidae"])
```

model - leech predictors
```{r}
fel_m1 <- glmmTMB(Number.of.Animals ~
         z_picta_part +
         z_zeylanica_part +
         samplingperiod +
         on.off.trail +
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$Family == "Felidae",],
         family = nbinom2,
         ziformula = ~1)

summary(fel_m1)

ggcoefstats(fel_m1, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  scale_y_discrete(labels = c("tiger leeches", "brown leeches", "sampling period", "on/off trail")) +
  labs(title = "Felidae - Leech Predictors Only")
```

model - leech counts plus ecological predictors 
```{r}
fel_m2 <- glmmTMB(Number.of.Animals ~
         z_picta_part +
         z_zeylanica_part +
         samplingperiod +
         on.off.trail +
         z_rain +                  
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$Family == "Felidae",],
         family = nbinom2,
         ziformula = ~1)

summary(fel_m2)
check_collinearity(fel_m2)

ggcoefstats(fel_m2, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  scale_y_discrete(labels = c("tiger leeches", "brown leeches", "sampling period", "on/off trail",
                              "rainfall", "max temp", "min temp", "fruit")) +
  labs(title = "Felidae - All Predictors")
```

model - full with interactions  
```{r}
fel_m3 <- glmmTMB(Number.of.Animals ~
         z_picta_part * z_rain +
         z_zeylanica_part * z_rain +
         samplingperiod +
         on.off.trail +
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$Family == "Felidae",],
         family = nbinom2,
         ziformula = ~1)

summary(fel_m3)
check_collinearity(fel_m3)

ggcoefstats(fel_m3, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  labs(title = "Felidae - All Predictors")
```

ecological only
```{r}
fel_m4 <- glmmTMB(Number.of.Animals ~
         samplingperiod +
         on.off.trail +
         z_rain +                  
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$Family == "Felidae",],
         family = nbinom2,
         ziformula = ~1)
```

model comparison
```{r}
AIC(fel_m1, fel_m2, fel_m3, fel_m4)
```

#### Hystricidae

Two species of porcupines, 1,426 total individuals, nearly even between the two species
```{r}
#n observations by species and by family
table(pruned$Species[pruned$Family == "Hystricidae" & pruned$num_observed != 0])
length(pruned$Species[pruned$Family == "Hystricidae" & pruned$num_observed != 0])

table(pruned$habitat[pruned$Family == "Hystricidae" & pruned$num_observed != 0])
table(pruned$partition[pruned$Family == "Hystricidae" & pruned$num_observed != 0])

hist(pruned$num_observed[pruned$Family == "Hystricidae"])
table(pruned$num_observed[pruned$Family == "Hystricidae"])
```

model - leech predictors
```{r}
hyst_m1 <- glmmTMB(Number.of.Animals ~
         z_picta_part +
         z_zeylanica_part +
         samplingperiod +
         on.off.trail +
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$Family == "Hystricidae",],
         family = nbinom2,
         ziformula = ~1)

summary(hyst_m1)

ggcoefstats(hyst_m1, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  scale_y_discrete(labels = c("tiger leeches", "brown leeches", "sampling period", "on/off trail")) +
  labs(title = "Hystricidae - Leech Predictors Only")
```

model - leech counts plus ecological predictors
```{r}
hyst_m2 <- glmmTMB(Number.of.Animals ~
         z_picta_part +
         z_zeylanica_part +
         samplingperiod +
         on.off.trail +
         z_rain +                  
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$Family == "Hystricidae",],
         family = nbinom2,
         ziformula = ~1)

summary(hyst_m2)
check_collinearity(hyst_m2)

ggcoefstats(hyst_m2, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  scale_y_discrete(labels = c("tiger leeches", "brown leeches", "sampling period", "on/off trail",
                              "rainfall", "max temp", "min temp", "fruit")) +
  labs(title = "Hystricidae - All Predictors")
```

model - full with interactions
```{r}
hyst_m3 <- glmmTMB(Number.of.Animals ~
         z_picta_part * z_rain +
         z_zeylanica_part * z_rain +
         samplingperiod +
         on.off.trail +
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$Family == "Hystricidae",],
         family = nbinom2,
         ziformula = ~1)

summary(hyst_m3)
check_collinearity(hyst_m3)

ggcoefstats(hyst_m3, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  labs(title = "Hystricidae - All Predictors with interactions")
```

eco only
```{r}
hyst_m4 <- glmmTMB(Number.of.Animals ~
         samplingperiod +
         on.off.trail +
         z_rain +                  
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$Family == "Hystricidae",],
         family = nbinom2,
         ziformula = ~1)
```

model comparison
```{r}
AIC(hyst_m1, hyst_m2, hyst_m3, hyst_m4)
```

#### Mustelidae

Three species of mustelids, almost all Martes flavigula
```{r}
#n observations by species and by family
table(pruned$Species[pruned$Family == "Mustelidae" & pruned$num_observed != 0])
length(pruned$Species[pruned$Family == "Mustelidae" & pruned$num_observed != 0])

table(pruned$habitat[pruned$Family == "Mustelidae" & pruned$num_observed != 0])
table(pruned$partition[pruned$Family == "Mustelidae" & pruned$num_observed != 0])

hist(pruned$num_observed[pruned$Family == "Mustelidae"])
table(pruned$num_observed[pruned$Family == "Mustelidae"])
```

model - leech predictors
```{r}
must_m1 <- glmmTMB(Number.of.Animals ~
         z_picta_part +
         z_zeylanica_part +
         on.off.trail +
         samplingperiod +       
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$Family == "Mustelidae",],
         family = nbinom2,
         ziformula = ~1)

summary(must_m1)

ggcoefstats(must_m1, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  scale_y_discrete(labels = c("tiger leeches", "brown leeches", "sampling period", "on/off trail")) +
  labs(title = "Mustelidae - Leech Predictors Only")
```

model - leech counts plus ecological predictors. No reliable predictors.
```{r}
must_m2 <- glmmTMB(Number.of.Animals ~
         z_picta_part +
         z_zeylanica_part +
         samplingperiod +
         on.off.trail +
         z_rain +                  
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod) +
         offset(log(active.days)),
         data = pruned_aggregated[pruned_aggregated$Family == "Mustelidae",],
         family = nbinom2,
         ziformula = ~1)

summary(must_m2)
check_collinearity(must_m2)

ggcoefstats(must_m2, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  scale_y_discrete(labels = c("tiger leeches", "brown leeches", "sampling period", "on/off trail",
                              "rainfall", "max temp", "min temp", "fruit")) +
  labs(title = "Mustelidae - All Predictors")
```

model - full with interactions
```{r}
must_m3 <- glmmTMB(Number.of.Animals ~
         z_picta_part * z_rain +
         z_zeylanica_part * z_rain +
         samplingperiod +
         on.off.trail +
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod) +
         offset(log(active.days)),
         data = pruned_aggregated[pruned_aggregated$Family == "Mustelidae",],
         family = nbinom2,
         ziformula = ~1)

summary(must_m3)
check_collinearity(must_m3)

ggcoefstats(must_m3, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  labs(title = "Mustelidae - All Predictors w/ interactions")
```

eco only
```{r}
must_m4 <- glmmTMB(Number.of.Animals ~
         samplingperiod +
         on.off.trail +
         z_rain +                  
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod) +
         offset(log(active.days)),
         data = pruned_aggregated[pruned_aggregated$Family == "Mustelidae",],
         family = nbinom2,
         ziformula = ~1)
```

model comparison
```{r}
AIC(must_m1, must_m2, must_m3, must_m4)
```


#### Sciuridae

Ten species of squirels, some Unid, most are Lariscus insignis 
```{r}
#n observations by species and by family
table(pruned$Species[pruned$Family == "Sciuridae" & pruned$num_observed != 0])
length(pruned$Species[pruned$Family == "Sciuridae" & pruned$num_observed != 0])

table(pruned$habitat[pruned$Family == "Sciuridae" & pruned$num_observed != 0])
table(pruned$partition[pruned$Family == "Sciuridae" & pruned$num_observed != 0])

hist(pruned$num_observed[pruned$Family == "Sciuridae"])
table(pruned$num_observed[pruned$Family == "Sciuridae"])
```

model - leech predictors
```{r}
sciu_m1 <- glmmTMB(Number.of.Animals ~
         z_picta_part +
         z_zeylanica_part +
         samplingperiod + 
         on.off.trail +   
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$Family == "Sciuridae",],
         family = nbinom2,
         ziformula = ~1)

summary(sciu_m1)

ggcoefstats(sciu_m1, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  scale_y_discrete(labels = c("tiger leeches", "brown leeches", "sampling period", "on/off trail")) +
  labs(title = "Sciuridae - Leech Predictors Only")
```

model - leech counts plus ecological predictors
```{r}
sciu_m2 <- glmmTMB(Number.of.Animals ~
         z_picta_part +
         z_zeylanica_part +
         samplingperiod +
         on.off.trail +
         z_rain +                  
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod) +
         offset(log(active.days)),
         data = pruned_aggregated[pruned_aggregated$Family == "Sciuridae",],
         family = nbinom2,
         ziformula = ~1)

summary(sciu_m2)
check_collinearity(sciu_m2)

ggcoefstats(sciu_m2, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  scale_y_discrete(labels = c("tiger leeches", "brown leeches", "sampling period", "on/off trail",
                              "rainfall", "max temp", "min temp", "fruit")) +
  labs(title = "Sciuridae - All Predictors")
```

model - full model with interactions
```{r}
sciu_m3 <- glmmTMB(Number.of.Animals ~
         z_picta_part * z_rain +
         z_zeylanica_part * z_rain +
         samplingperiod +
         on.off.trail +
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod) +
         offset(log(active.days)),
         data = pruned_aggregated[pruned_aggregated$Family == "Sciuridae",],
         family = nbinom2,
         ziformula = ~1)

summary(sciu_m3)
check_collinearity(sciu_m3)

ggcoefstats(sciu_m3, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  labs(title = "Sciuridae - All Predictors w/ interactions")
```

interaction plot
```{r}
p1 <- plot_model(sciu_m3, type = "pred", terms = c("z_picta_part", "z_rain [minmax]")) +
  theme_classic() +
  labs(x = "tiger leech rate (z)", y = "mammal count", 
       title = "picta X rainfall", color = "rainfall (z)") +
  coord_cartesian(ylim = c(0, 3)) +
  theme(legend.position = "bottom")

p2 <- plot_model(sciu_m3, type = "pred", terms = c("z_zeylanica_part", "z_rain [minmax]")) +
  theme_classic() +
  labs(x = "brown leech rate (z)", y = "mammal count", 
       title = "zeylanica X rainfall", color = "rainfall (z)") +
  coord_cartesian(ylim = c(0, 3), xlim = c(-1, 1.5)) +
  theme(legend.position = "bottom")

grid.arrange(p1, p2, nrow = 1)
```

eco only
```{r}
sciu_m4 <- glmmTMB(Number.of.Animals ~
         samplingperiod +
         on.off.trail +
         z_rain +                  
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod) +
         offset(log(active.days)),
         data = pruned_aggregated[pruned_aggregated$Family == "Sciuridae",],
         family = nbinom2,
         ziformula = ~1)
```

model comparison
```{r}
AIC(sciu_m1, sciu_m2, sciu_m3, sciu_m4)
```


#### Tragulidae

Two species of Tragulidae, mostly T. javanicus
```{r}
#n observations by species and by family
table(pruned$Species[pruned$Family == "Tragulidae" & pruned$num_observed != 0])
length(pruned$Species[pruned$Family == "Tragulidae" & pruned$num_observed != 0])

table(pruned$habitat[pruned$Family == "Tragulidae" & pruned$num_observed != 0])
table(pruned$partition[pruned$Family == "Tragulidae" & pruned$num_observed != 0])

hist(pruned$num_observed[pruned$Family == "Tragulidae"])
table(pruned$num_observed[pruned$Family == "Tragulidae"])
```

model - leech predictors
```{r}
trag_m1 <- glmmTMB(Number.of.Animals ~
         z_picta_part +
         z_zeylanica_part +
         samplingperiod +
         on.off.trail +
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$Family == "Tragulidae",],
         family = nbinom2,
         ziformula = ~1)

summary(trag_m1)

ggcoefstats(trag_m1, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  scale_y_discrete(labels = c("tiger leeches", "brown leeches", "sampling period", "on/off trail")) +
  labs(title = "Tragulidae - Leech Predictors Only")
```

model - leech counts plus ecological predictors
```{r}
trag_m2 <- glmmTMB(Number.of.Animals ~
         z_picta_part +
         z_zeylanica_part +
         samplingperiod +
         on.off.trail +
         z_rain +                  
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$Family == "Tragulidae",],
         family = nbinom2,
         ziformula = ~1)

summary(trag_m2)
check_collinearity(trag_m2)

ggcoefstats(trag_m2, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  scale_y_discrete(labels = c("tiger leeches", "brown leeches", "sampling period", "on/off trail",
                              "rainfall", "max temp", "min temp", "fruit")) +
  labs(title = "Tragulidae - All Predictors")
```

model - full model with interactions
```{r}
trag_m3 <- glmmTMB(Number.of.Animals ~
         z_picta_part * z_rain +
         #z_zeylanica_part * z_rain + #high correlation, VIF > 14
         samplingperiod +
         on.off.trail +
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$Family == "Tragulidae",],
         family = nbinom2,
         ziformula = ~1)

summary(trag_m3)
check_collinearity(trag_m3)

ggcoefstats(trag_m3, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  labs(title = "Tragulidae - All Predictors w/ interactions")
```

interaction plot
```{r}
p1 <- plot_model(trag_m3, type = "pred", terms = c("z_picta_part", "z_rain [minmax]")) +
  theme_classic() +
  labs(x = "tiger leech rate (z)", y = "mammal count", 
       title = "picta X rainfall", color = "rainfall (z)") +
  coord_cartesian(ylim = c(0, 0.5)) +
  theme(legend.position = "bottom")
```


eco only
```{r}
trag_m4 <- glmmTMB(Number.of.Animals ~
         samplingperiod +
         on.off.trail +
         z_rain +                  
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$Family == "Tragulidae",],
         family = nbinom2,
         ziformula = ~1)
```

model comparison
```{r}
AIC(trag_m1, trag_m2, trag_m3, trag_m4)
```

interaction plot
```{r}
plot_model(trag_m3, type = "pred", terms = c("z_picta_part", "z_rain [minmax]")) +
  theme_classic() +
  labs(x = "tiger leech rate (z)", y = "mammal count", 
       title = "leech X rainfall interaction", color = "rainfall (z)")
```

#### Viverridae

Eight species of Viverridae, most are V.tangalunga
```{r}
#n observations by species and by family
table(pruned$Species[pruned$Family == "Viverridae" & pruned$num_observed != 0])
length(pruned$Species[pruned$Family == "Viverridae" & pruned$num_observed != 0])

table(pruned$habitat[pruned$Family == "Viverridae" & pruned$num_observed != 0])
table(pruned$partition[pruned$Family == "Viverridae" & pruned$num_observed != 0])

hist(pruned$num_observed[pruned$Family == "Viverridae"])
table(pruned$num_observed[pruned$Family == "Viverridae"])
```

model - leech predictors
```{r}
viv_m1 <- glmmTMB(Number.of.Animals ~
         z_picta_part +
         z_zeylanica_part +
         samplingperiod + 
         on.off.trail +
         (1|partition.mod),
         offset =log(active.days),
         data = pruned_aggregated[pruned_aggregated$Family == "Viverridae",],
         family = nbinom2,
         ziformula = ~1)

summary(viv_m1)

ggcoefstats(viv_m1, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  scale_y_discrete(labels = c("tiger leeches", "brown leeches", "sampling period", "on/off trail")) +
  labs(title = "Viverridae - Leech Predictors Only")
```

model - leech counts plus ecological predictors
```{r}
viv_m2 <- glmmTMB(Number.of.Animals ~
         z_picta_part +
         z_zeylanica_part +
         samplingperiod +
         on.off.trail +
         z_rain +                  
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod) +
         offset(log(active.days)),
         data = pruned_aggregated[pruned_aggregated$Family == "Viverridae",],
         family = nbinom2,
         ziformula = ~1)

summary(viv_m2)
check_collinearity(viv_m2)

ggcoefstats(viv_m2, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  scale_y_discrete(labels = c("tiger leeches", "brown leeches", "sampling period", "on/off trail",
                              "rainfall", "max temp", "min temp", "fruit")) +
  labs(title = "Viverridae - All Predictors")
```

model - full with interactions
```{r}
viv_m3 <- glmmTMB(Number.of.Animals ~
         z_picta_part * z_rain +
         z_zeylanica_part * z_rain +
         samplingperiod +
         on.off.trail +
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod) +
         offset(log(active.days)),
         data = pruned_aggregated[pruned_aggregated$Family == "Viverridae",],
         family = nbinom2,
         ziformula = ~1)

summary(viv_m3)
check_collinearity(viv_m3)

ggcoefstats(viv_m3, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  labs(title = "Viverridae - All Predictors with interactions")
```

eco only
```{r}
viv_m4 <- glmmTMB(Number.of.Animals ~
         samplingperiod +
         on.off.trail +
         z_rain +                  
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod) +
         offset(log(active.days)),
         data = pruned_aggregated[pruned_aggregated$Family == "Viverridae",],
         family = nbinom2,
         ziformula = ~1)
```

model comparison
```{r}
AIC(viv_m1, viv_m2, viv_m3, viv_m4)
```


### Grouping by functional groups

#### Activity Cycle - Nocturnal

21 species, 7,559 observations
```{r}
#observations by species
table(pruned$Species[pruned$ActivityCycle == 1 & pruned$num_observed != 0])
length(pruned$Species[pruned$ActivityCycle == 1 & pruned$num_observed != 0])

table(pruned$habitat[pruned$ActivityCycle == 1 & pruned$num_observed != 0])
table(pruned$partition[pruned$ActivityCycle == 1 & pruned$num_observed != 0])

hist(pruned$num_observed[pruned$ActivityCycle == 1])
table(pruned$num_observed[pruned$ActivityCycle == 1])
```

model with just leech predictors
```{r}
noct_m1 <- glmmTMB(Number.of.Animals ~
         z_picta_part +
         z_zeylanica_part +
         samplingperiod +
         on.off.trail +
         (1|partition.mod) +
         offset(log(active.days)),
         data = pruned_aggregated[pruned_aggregated$ActivityCycle == 1,],
         family = nbinom2,
         ziformula = ~1)

summary(noct_m1)

ggcoefstats(noct_m1, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  scale_y_discrete(labels = c("tiger leeches", "brown leeches", "sampling period", "on/off trail")) +
  labs(title = "Nocturnal Vertebrates - Leech Predictors Only")
```

model - leech counts plus ecological predictors 
```{r}
noct_m2 <- glmmTMB(Number.of.Animals ~
         z_picta_part +
         z_zeylanica_part +
         samplingperiod +
         on.off.trail +
         z_rain +                  
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod) +
         offset(log(active.days)),
         data = pruned_aggregated[pruned_aggregated$ActivityCycle == 1,],
         family = nbinom2,
         ziformula = ~1)

summary(noct_m2)
check_collinearity(noct_m2)

ggcoefstats(noct_m2, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  scale_y_discrete(labels = c("tiger leeches", "brown leeches", "sampling period", "on/off trail",
                              "rainfall", "max temp", "min temp", "fruit")) +
  labs(title = "Nocturnal Vertebrates - All Predictors")
```

model - full model with interactions
```{r}
noct_m3 <- glmmTMB(Number.of.Animals ~
         z_picta_part * z_rain +
         z_zeylanica_part * z_rain +
         samplingperiod +
         on.off.trail +
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod) +
         offset(log(active.days)),
         data = pruned_aggregated[pruned_aggregated$ActivityCycle == 1,],
         family = nbinom2,
         ziformula = ~1)

summary(noct_m3)
check_collinearity(noct_m3)

ggcoefstats(noct_m3, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  labs(title = "Nocturnal Vertebrates - All Predictors with interactions")
```

eco only 
```{r}
noct_m4 <- glmmTMB(Number.of.Animals ~
         samplingperiod +
         on.off.trail +
         z_rain +                  
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod) +
         offset(log(active.days)),
         data = pruned_aggregated[pruned_aggregated$ActivityCycle == 1,],
         family = nbinom2,
         ziformula = ~1)
```

model comparison
```{r}
AIC(noct_m1, noct_m2, noct_m3, noct_m4)
```

#### Activity Cycle - Crepuscular

8 species, 4,029 observations. 56% are Sus barbatus
```{r}
#observations by species
table(pruned$Species[pruned$ActivityCycle == 2 & pruned$num_observed != 0])
length(pruned$Species[pruned$ActivityCycle == 2 & pruned$num_observed != 0])

table(pruned$habitat[pruned$ActivityCycle == 2 & pruned$num_observed != 0])
table(pruned$partition[pruned$ActivityCycle == 2 & pruned$num_observed != 0])

hist(pruned$num_observed[pruned$ActivityCycle == 2])
table(pruned$num_observed[pruned$ActivityCycle == 2])
```

model with just leech predictors - both leeches are reliable predictors 
```{r}
crep_m1 <- glmmTMB(Number.of.Animals ~
         z_picta_part +
         z_zeylanica_part +
         samplingperiod + 
         on.off.trail +
         (1|partition.mod) +
         offset(log(active.days)),
         data = pruned_aggregated[pruned_aggregated$ActivityCycle == 2,],
         family = nbinom2,
         ziformula = ~1)

summary(crep_m1)

ggcoefstats(crep_m1, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  scale_y_discrete(labels = c("tiger leeches", "brown leeches", "sampling period", "on/off trail")) +
  labs(title = "Crepuscular Vertebrates - Leech Predictors Only")
```

model - leech counts plus ecological predictors.
```{r}
crep_m2 <- glmmTMB(Number.of.Animals ~
         z_picta_part +
         z_zeylanica_part +
         samplingperiod +
         on.off.trail +
         z_rain +                  
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$ActivityCycle == 2,],
         family = nbinom2,
         ziformula = ~1)

summary(crep_m2)
check_collinearity(crep_m2)

ggcoefstats(crep_m2, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  scale_y_discrete(labels = c("tiger leeches", "brown leeches", "sampling period", "on/off trail",
                              "rainfall", "max temp", "min temp", "fruit")) +
  labs(title = "Crepuscular Vertebrates - All Predictors")
```

model - full model with interactions
```{r}
crep_m3 <- glmmTMB(Number.of.Animals ~
         z_picta_part * z_rain +
         z_zeylanica_part * z_rain +
         samplingperiod +
         on.off.trail +
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$ActivityCycle == 2,],
         family = nbinom2,
         ziformula = ~1)

summary(crep_m3)
check_collinearity(crep_m3)

ggcoefstats(crep_m3, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  labs(title = "Crepuscular Vertebrates - All Predictors with interactions")
```

eco only
```{r}
crep_m4 <- glmmTMB(Number.of.Animals ~
         samplingperiod +
         on.off.trail +
         z_rain +                  
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$ActivityCycle == 2,],
         family = nbinom2,
         ziformula = ~1)
```

model comparison
```{r}
AIC(crep_m1, crep_m2, crep_m3, crep_m4)
```


#### Activity Cycle - Diurnal

23 species, 3,044 observations. 31% are M.nemestrina
```{r}
#observations by species
table(pruned$Species[pruned$ActivityCycle == 3 & pruned$num_observed != 0])
length(pruned$Species[pruned$ActivityCycle == 3 & pruned$num_observed != 0])

table(pruned$habitat[pruned$ActivityCycle == 3 & pruned$num_observed != 0])
table(pruned$partition[pruned$ActivityCycle == 3 & pruned$num_observed != 0])

hist(pruned$num_observed[pruned$ActivityCycle == 3])
table(pruned$num_observed[pruned$ActivityCycle == 3])
```

model with just leech predictors
```{r}
diur_m1 <- glmmTMB(Number.of.Animals ~
         z_picta_part +
         z_zeylanica_part +
         samplingperiod + 
         on.off.trail +
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$ActivityCycle == 3,],
         family = nbinom2,
         ziformula = ~1)

summary(diur_m1)

ggcoefstats(diur_m1, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  scale_y_discrete(labels = c("tiger leeches", "brown leeches", "sampling period", "on/off trail")) +
  labs(title = "Diurnal Vertebrates - Leech Predictors Only")
```

model - leech counts plus ecological predictors.
```{r}
diur_m2 <- glmmTMB(Number.of.Animals ~
         z_picta_part +
         z_zeylanica_part +
         samplingperiod +
         on.off.trail +
         z_rain +                  
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$ActivityCycle == 3,],
         family = nbinom2,
         ziformula = ~1)

summary(diur_m2)
check_collinearity(diur_m2)

ggcoefstats(diur_m2, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  scale_y_discrete(labels = c("tiger leeches", "brown leeches", "sampling period", "on/off trail",
                              "rainfall", "max temp", "min temp", "fruit")) +
  labs(title = "Diurnal Vertebrates - All Predictors")
```

model - full model with interactions
```{r}
diur_m3 <- glmmTMB(Number.of.Animals ~
         z_picta_part * z_rain +
         z_zeylanica_part * z_rain +
         samplingperiod +
         on.off.trail +
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$ActivityCycle == 3,],
         family = nbinom2,
         ziformula = ~1)

summary(diur_m3)
check_collinearity(diur_m3)

ggcoefstats(diur_m3, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  labs(title = "Diurnal Vertebrates - All Predictors with interactions")
```

eco only
```{r}
diur_m4 <- glmmTMB(Number.of.Animals ~
         samplingperiod +
         on.off.trail +
         z_rain +                  
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$ActivityCycle == 3,],
         family = nbinom2,
         ziformula = ~1)
```

model comparison
```{r}
AIC(diur_m1, diur_m2, diur_m3, diur_m4)
```


#### Trophic Level - Herbivore

13 species, 5,598 observations
```{r}
#observations by species
table(pruned$Species[pruned$TrophicLevel == 1])
length(pruned$Species[pruned$TrophicLevel == 1])

table(pruned$habitat[pruned$TrophicLevel == 1])
table(pruned$partition[pruned$TrophicLevel == 1])

hist(pruned$num_observed[pruned$TrophicLevel == 1])
table(pruned$num_observed[pruned$TrophicLevel == 1])
```

model with just leech predictors
```{r}
herb_m1 <- glmmTMB(Number.of.Animals ~
         z_picta_part +
         z_zeylanica_part +
         samplingperiod +
         on.off.trail +
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$TrophicLevel == 1,],
         family = nbinom2,
         ziformula = ~1)

summary(herb_m1)

ggcoefstats(herb_m1, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  scale_y_discrete(labels = c("tiger leeches", "brown leeches", "sampling period", "on/off trail")) +
  labs(title = "Herbivorous Vertebrates - Leech Predictors Only")
```

model - leech counts plus ecological predictors
```{r}
herb_m2 <- glmmTMB(Number.of.Animals ~
         z_picta_part +
         z_zeylanica_part +
         samplingperiod +
         on.off.trail +
         z_rain +                  
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod) +
         offset(log(active.days)),
         data = pruned_aggregated[pruned_aggregated$TrophicLevel == 1,],
         family = nbinom2,
         ziformula = ~1)

summary(herb_m2)
check_collinearity(herb_m2)

ggcoefstats(herb_m2, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  scale_y_discrete(labels = c("tiger leeches", "brown leeches", "sampling period", "on/off trail",
                              "rainfall", "max temp", "min temp", "fruit")) +
  labs(title = "Herbivorous Vertebrates - All Predictors")
```

model - full plus interactions
```{r}
herb_m3 <- glmmTMB(Number.of.Animals ~
         z_picta_part * z_rain +
         z_zeylanica_part * z_rain +
         samplingperiod +
         on.off.trail +
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod) +
         offset(log(active.days)),
         data = pruned_aggregated[pruned_aggregated$TrophicLevel == 1,],
         family = nbinom2,
         ziformula = ~1)

summary(herb_m3)
check_collinearity(herb_m3)

ggcoefstats(herb_m3, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  labs(title = "Herbivorous Vertebrates - All Predictors with interactions")
```

eco only
```{r}
herb_m4 <- glmmTMB(Number.of.Animals ~
         samplingperiod +
         on.off.trail +
         z_rain +                  
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod) +
         offset(log(active.days)),
         data = pruned_aggregated[pruned_aggregated$TrophicLevel == 1,],
         family = nbinom2,
         ziformula = ~1)
```

model comparison
```{r}
AIC(herb_m1, herb_m2, herb_m3, herb_m4)
```


#### Trophic Level - Omnivore

30 species, 9,510 observations
```{r}
#observations by species
table(pruned$Species[pruned$TrophicLevel == 2])
length(pruned$Species[pruned$TrophicLevel == 2])

table(pruned$habitat[pruned$TrophicLevel == 2])
table(pruned$partition[pruned$TrophicLevel == 2])

hist(pruned$num_observed[pruned$TrophicLevel == 2])
table(pruned$num_observed[pruned$TrophicLevel == 2])
```

model with just leech predictors
```{r}
omni_m1 <- glmmTMB(Number.of.Animals ~
         z_picta_part +
         z_zeylanica_part +
         samplingperiod +
         on.off.trail +
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$TrophicLevel == 2,],
         family = nbinom2,
         ziformula = ~1)

summary(omni_m1)

ggcoefstats(omni_m1, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  scale_y_discrete(labels = c("tiger leeches", "brown leeches", "sampling period", "on/off trail")) +
  labs(title = "Omnivorous Vertebrates - Leech Predictors Only")
```

model - leech counts plus ecological predictors.
```{r}
omni_m2 <- glmmTMB(Number.of.Animals ~
         z_picta_part +
         z_zeylanica_part +
         samplingperiod +
         on.off.trail +
         z_rain +                  
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod) +
         offset(log(active.days)),
         data = pruned_aggregated[pruned_aggregated$TrophicLevel == 2,],
         family = nbinom2,
         ziformula = ~1)

summary(omni_m2)
check_collinearity(omni_m2)

ggcoefstats(omni_m2, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  scale_y_discrete(labels = c("tiger leeches", "brown leeches", "sampling period", "on/off trail",
                              "rainfall", "max temp", "min temp", "fruit")) +
  labs(title = "Omnivorous Vertebrates - All Predictors")
```

model - full model with interactions
```{r}
omni_m3 <- glmmTMB(Number.of.Animals ~
         z_picta_part * z_rain +
         z_zeylanica_part * z_rain +
         samplingperiod +
         on.off.trail +
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod) +
         offset(log(active.days)),
         data = pruned_aggregated[pruned_aggregated$TrophicLevel == 2,],
         family = nbinom2,
         ziformula = ~1)

summary(omni_m3)
check_collinearity(omni_m3)

ggcoefstats(omni_m3, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  labs(title = "Omnivorous Vertebrates - All Predictors")
```

eco only
```{r}
omni_m4 <- glmmTMB(Number.of.Animals ~
         samplingperiod +
         on.off.trail +
         z_rain +                  
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod) +
         offset(log(active.days)),
         data = pruned_aggregated[pruned_aggregated$TrophicLevel == 2,],
         family = nbinom2,
         ziformula = ~1)
```

model comparison
```{r}
AIC(omni_m1, omni_m2, omni_m3, omni_m4)
```


#### Trophic Level - Carnivore

9 species, 1,231 observations
```{r}
#observations by species
table(pruned$Species[pruned$TrophicLevel == 3])
length(pruned$Species[pruned$TrophicLevel == 3])

table(pruned$habitat[pruned$TrophicLevel == 3])
table(pruned$partition[pruned$TrophicLevel == 3])

hist(pruned$num_observed[pruned$TrophicLevel == 3])
table(pruned$num_observed[pruned$TrophicLevel == 3])
```

model with just leech predictors
```{r}
carn_m1 <- glmmTMB(Number.of.Animals ~
         z_picta_part +
         z_zeylanica_part +
         samplingperiod +  
         on.off.trail +
         (1|partition.mod),
         offset = log(active.days),
         data = pruned_aggregated[pruned_aggregated$TrophicLevel == 3,],
         family = nbinom2,
         ziformula = ~1)

summary(carn_m1)

ggcoefstats(carn_m1, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  scale_y_discrete(labels = c("tiger leeches", "brown leeches", "sampling period", "on/off trail")) +
  labs(title = "Carnivorous Vertebrates - Leech Predictors Only")
```

model - leech counts plus ecological predictors.
```{r}
carn_m2 <- glmmTMB(Number.of.Animals ~
         z_picta_part +
         z_zeylanica_part +
         samplingperiod +
         on.off.trail +
         z_rain +                  
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod) +
         offset(log(active.days)),
         data = pruned_aggregated[pruned_aggregated$TrophicLevel == 3,],
         family = nbinom2,
         ziformula = ~1)

summary(carn_m2)
check_collinearity(carn_m2)

ggcoefstats(carn_m2, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  scale_y_discrete(labels = c("tiger leeches", "brown leeches", "sampling period", "on/off trail",
                              "rainfall", "max temp", "min temp", "fruit")) +
  labs(title = "Carnivorous Vertebrates - All Predictors")
```

model - full model with interactions
```{r}
carn_m3 <- glmmTMB(Number.of.Animals ~
         z_picta_part * z_rain +
         z_zeylanica_part * z_rain +
         samplingperiod +
         on.off.trail +
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod) +
         offset(log(active.days)),
         data = pruned_aggregated[pruned_aggregated$TrophicLevel == 3,],
         family = nbinom2,
         ziformula = ~1)

summary(carn_m3)
check_collinearity(carn_m3)

ggcoefstats(carn_m3, exclude.intercept = TRUE, stats.labels = FALSE, ylab = "") +
  theme_classic() +
  labs(title = "Carnivorous Vertebrates - All Predictors")
```

```{r}
carn_m4 <- glmmTMB(Number.of.Animals ~
         samplingperiod +
         on.off.trail +
         z_rain +                  
         z_max +                   
         z_min +                   
         z_pheno_mod +                 
         (1|partition.mod) +
         offset(log(active.days)),
         data = pruned_aggregated[pruned_aggregated$TrophicLevel == 3,],
         family = nbinom2,
         ziformula = ~1)
```

model comparison
```{r}
AIC(carn_m1, carn_m2, carn_m3, carn_m4)
```


## Visualizing effect of leeches across models

#### full models

```{r}
#model list - full models
model_list <- list(mod2, rats_m2, pigs_m2, vtan_m2, tjav_m2, mnem_m2, cerv_m2, fel_m2, hyst_m2, must_m2, sciu_m2, trag_m2, viv_m2, noct_m2, crep_m2, diur_m2, herb_m2, omni_m2, carn_m2)

#extract leech coefficients
coef_list <- lapply(model_list, function(model) {
  as.data.frame(summary(model)$coefficients$cond[c("z_zeylanica_part", "z_picta_part"), ])
})

#combine coefs into data frame
coef_df <- do.call(rbind, coef_list)

#add model names
coef_df$model <- rep(c("All Mammals", "Unidentified Rats", "Bearded Pig", "Malay Civet", "Mouse Deer", "Pig-tailed Macaque",  "Cervidae",  "Felidae", "Hystricidae", "Mustelidae", "Sciuridae", "Tragulidae", "Viverridae", "Nocturnal", "Crepuscular", "Diurnal", "Herbivores", "Omnivores", "Carnivores"), each = 2)

#add leech names
coef_df$predictor <- rep(c("brown leech", "tiger leech"), times = length(model_list))

#add confidence intervals
coef_df$conf.low <- coef_df$Estimate - 1.96 * coef_df$`Std. Error`
coef_df$conf.high <- coef_df$Estimate + 1.96 * coef_df$`Std. Error`
coef_df$conf.low_50 <- coef_df$Estimate - 0.674 * coef_df$`Std. Error`
coef_df$conf.high_50 <- coef_df$Estimate + 0.674 * coef_df$`Std. Error`


#plot
coef_df %>%
  mutate(model = factor(model, levels = coef_df %>%
                          filter(predictor == "tiger leech") %>%
                          arrange(Estimate) %>%
                          pull(model))) %>%
ggplot(aes(x = model, y = Estimate, color = predictor)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0, 
                position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(ymin = conf.low_50, ymax = conf.high_50), width = 0, 
                position = position_dodge(width = 0.5), size = 1.1) +
  theme_classic() +
  labs(title = "", y = "coefficient estimate") +
  coord_flip() +  
  scale_color_manual(values = c("tiger leech" = "orange", "brown leech" = "brown")) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey60") +  
  theme(axis.title.y = element_blank(),  
        legend.title = element_blank(),
        legend.position = c(0.85, 0.45)) +
  guides(color = guide_legend(reverse = TRUE))

#ggsave(filename = "LeechEffects_Multiplot.jpg", width = 4, height = 5, dpi = 350)
```


### Rainfall threshold models
```{r}
#function to refit models using different rain thresholds
refit_model_with_rain_filter <- function(model, data, threshold = "all") {
  cutoff <- switch(threshold,
                   "top50" = quantile(data$z_rain, 0.5, na.rm = TRUE),  #top half rainy periods
                   "top25" = quantile(data$z_rain, 0.75, na.rm = TRUE), #top quarter rainy periods
                   -Inf) 
  
  filtered_data <- if (threshold == "all") data else filter(data, z_rain > cutoff)
  
  update(model, data = filtered_data)
}

#add model names
names(model_list) <- c("All Mammals", "Rats", "Pigs", "Malay Civet", "Mouse Deer", "PT Macaque", "Cervidae", "Felidae", "Hystricidae", "Mustelidae", "Sciuridae", "Tragulidae", "Viverridae", "Nocturnal", "Crepuscular", "Diurnal", "Herbivores", "Omnivores", "Carnivores")

#create data list
data_list <- list(
  pruned_aggregated,
  pruned_aggregated[pruned_aggregated$Species == "Unid Rat",],
  pruned_aggregated[pruned_aggregated$Species == "Sus barbatus",],
  pruned_aggregated[pruned_aggregated$Species == "Viverra tangalunga",],
  pruned_aggregated[pruned_aggregated$Species == "Tragulus javanicus",],
  pruned_aggregated[pruned_aggregated$Species == "Macaca nemestrina",],
  pruned_aggregated[pruned_aggregated$Family == "Cervidae",],  
  pruned_aggregated[pruned_aggregated$Family == "Felidae",],
  pruned_aggregated[pruned_aggregated$Family == "Hystricidae",],
  pruned_aggregated[pruned_aggregated$Family == "Mustelidae",],
  pruned_aggregated[pruned_aggregated$Family == "Sciuridae",],
  pruned_aggregated[pruned_aggregated$Family == "Tragulidae",],
  pruned_aggregated[pruned_aggregated$Family == "Viverridae",],
  pruned_aggregated[pruned_aggregated$ActivityCycle == 1,],
  pruned_aggregated[pruned_aggregated$ActivityCycle == 2,],
  pruned_aggregated[pruned_aggregated$ActivityCycle == 3,],
  pruned_aggregated[pruned_aggregated$TrophicLevel == 1,],
  pruned_aggregated[pruned_aggregated$TrophicLevel == 2,],
  pruned_aggregated[pruned_aggregated$TrophicLevel == 3,])


#define predictors of interest
predictors_of_interest <- c("z_picta_part", "z_zeylanica_part")
thresholds <- c("all", "top50", "top25")
results_list <- list()

#loop over both models
for (i in seq_along(model_list)) {
  model <- model_list[[i]]
  species_name <- names(model_list)[i]
  data <- data_list[[i]]
  
  for (thresh in thresholds) {
    cutoff <- switch(thresh,
                     "top50" = quantile(data$z_rain, 0.5, na.rm = TRUE),
                     "top25" = quantile(data$z_rain, 0.75, na.rm = TRUE),
                     -Inf)
    
    filtered_data <- if (thresh == "all") data else dplyr::filter(data, z_rain > cutoff)
    
    mod_refit <- update(model, data = filtered_data)
    
    coefs <- broom.mixed::tidy(mod_refit, effects = "fixed") %>%
      filter(term %in% predictors_of_interest) %>%
      mutate(species = species_name, threshold = thresh)
    
    results_list[[paste(species_name, thresh, sep = "_")]] <- coefs
  }
}

#combine
coef_df <- dplyr::bind_rows(results_list)
coef_df$estimate

#add confidence intervals
coef_df$conf.low <- coef_df$estimate - 1.96 * coef_df$std.error
coef_df$conf.high <- coef_df$estimate + 1.96 * coef_df$std.error
coef_df$conf.low_50 <- coef_df$estimate - 0.674 * coef_df$std.error
coef_df$conf.high_50 <- coef_df$estimate + 0.674 * coef_df$std.error

#plot
coef_df %>%
  mutate(species = case_when(species == "PT Macaque" ~ "Pig-tailed Macaque", TRUE ~ species),
         species = case_when(species == "Pigs" ~ "Bearded Pig", TRUE ~ species),
         species = case_when(species == "Rats" ~ "Unidentified Rats", TRUE ~ species),
    threshold = factor(threshold, levels = c("all", "top50", "top25")),
    species = factor(species, levels = rev(unique(species)))) %>%
  ggplot(aes(x = species, y = estimate, color = threshold)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0, 
                position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(ymin = conf.low_50, ymax = conf.high_50), width = 0, 
                position = position_dodge(width = 0.5), size = 1.1) +
  coord_flip() + 
  facet_wrap(~ term, scales = "free_x",
             labeller = as_labeller(c(
  "z_picta_part" = "Effect of Tiger Leech",
  "z_zeylanica_part" = "Effect of Brown Leech"))) + 
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray60") +
  labs(x = "", y = "coefficient estimate", 
  title = "",
    color = "rainfall threshold") +
  theme_classic(base_size = 12) +
  theme(panel.grid.major.y = element_blank(),
        axis.text.y = element_text(size = 10),
        strip.background = element_blank(),
        legend.position = "bottom") +
  guides(color = guide_legend(reverse = TRUE))

#ggsave("RainThresholdPlot.jpg", width = 6, height = 7, dpi = 300)
```

AICc table comparing models at different rain levels: with leech predictors (full model) and without (leech only, eco vars only)
```{r eval=FALSE, include=FALSE}
#results list
aic_results <- list()

#loop through models
for (i in seq_along(model_list)) {
  model <- model_list[[i]]
  species_name <- names(model_list)[i]
  data <- data_list[[i]]
  
  aic_entry <- list(Species = species_name)
  
  for (thresh in thresholds) {
    cutoff <- switch(thresh,
                     "top50" = quantile(data$z_rain, 0.5, na.rm = TRUE),
                     "top25" = quantile(data$z_rain, 0.75, na.rm = TRUE),
                     -Inf)
    
    filtered_data <- if (thresh == "all") data else dplyr::filter(data, z_rain > cutoff)
    
    full_model <- update(model, data = filtered_data)
    full_aic <- AICc(full_model)
    
    reduced_formula <- update(formula(model), paste(". ~ . -", paste(predictors_of_interest, 
                                                                     collapse = " - ")))
    reduced_model <- glmmTMB::glmmTMB(formula = reduced_formula, data = filtered_data, 
                                      family = family(model))
    reduced_aic <- AICc(reduced_model)
    
    aic_entry[[paste0("AICc_", thresh)]] <- full_aic
    aic_entry[[paste0("AICc_noleech_", thresh)]] <- reduced_aic
  }
  
  aic_results[[species_name]] <- aic_entry
}

#make dataframe
aic_table <- do.call(rbind, lapply(aic_results, as.data.frame))
rownames(aic_table) <- NULL

#which model had the lower AIC for each threshold
aic_table$lower_all <- ifelse(aic_table$AIC_all < aic_table$AIC_noleech_all, "full", "no_leech")
aic_table$lower_top50 <- ifelse(aic_table$AIC_top50 < aic_table$AIC_noleech_top50, "full", "no_leech")
aic_table$lower_top25 <- ifelse(aic_table$AIC_top25 < aic_table$AIC_noleech_top25, "full", "no_leech")

aic_table <- aic_table[, c("Species", 
                           "AIC_all", "AIC_noleech_all", "lower_all",
                           "AIC_top50", "AIC_noleech_top50", "lower_top50",
                           "AIC_top25", "AIC_noleech_top25", "lower_top25")]

colnames(aic_table) <- c("Species", "all", "all_nl", "low_all", "50", 
                         "50_nl", "low_50", "25", "25_nl", "low_25")

aic_table[] <- lapply(aic_table, function(x) if (is.numeric(x)) round(x, 1) else x)

aic_table
```

again but fitting models minus one of each leech predictor in addition to both
```{r}
#define leech predictors 
zey_predictor <- "z_zeylanica_part"
picta_predictor <- "z_picta_part"
both_predictors <- c(zey_predictor, picta_predictor)

#results list
aic_results <- list()

#loop 
for (i in seq_along(model_list)) {
  model <- model_list[[i]]
  species_name <- names(model_list)[i]
  data <- data_list[[i]]
  
  aic_entry <- list(Species = species_name)
  
  for (thresh in thresholds) {
    #filter by rainfall threshold
    cutoff <- switch(thresh,
                     "top50" = quantile(data$z_rain, 0.5, na.rm = TRUE),
                     "top25" = quantile(data$z_rain, 0.75, na.rm = TRUE),
                     -Inf)
    
    filtered_data <- if (thresh == "all") data else dplyr::filter(data, z_rain > cutoff)
    response_var <- deparse(formula(model)[[2]])
    aic_vals <- list()
    
    #full model
    mod_full <- update(model, data = filtered_data)
    aic_vals$full <- AICc(mod_full)
    
    #no zeylanica
    #form_no_zey <- update(formula(model), paste(". ~ . -", zey_predictor))
    #mod_no_zey <- glmmTMB::glmmTMB(formula = form_no_zey, data = filtered_data, family = family(model))
    #aic_vals$no_zeylanica <- AICc(mod_no_zey)
    
    #no picta
    #form_no_picta <- update(formula(model), paste(". ~ . -", picta_predictor))
    #mod_no_picta <- glmmTMB::glmmTMB(formula = form_no_picta, data = filtered_data, family = family(model))
    #aic_vals$no_picta <- AICc(mod_no_picta)
    
    #no leeches - eco vars only
    form_no_leech <- update(formula(model), paste(". ~ . -", paste(both_predictors, collapse = " - ")))
    mod_no_leech <- glmmTMB::glmmTMB(formula = form_no_leech, data = filtered_data, family = family(model))
    aic_vals$no_leech <- AICc(mod_no_leech)
    
    #leech-only (picta + zeylanica)
    form_leech_only <- as.formula(paste(response_var, "~", paste(both_predictors, collapse = " + ")))
    mod_leech_only <- glmmTMB::glmmTMB(formula = form_leech_only, data = filtered_data, family = family(model))
    aic_vals$leech_only <- AICc(mod_leech_only)
    
    #zeylanica-only
    #form_zey_only <- as.formula(paste(response_var, "~", zey_predictor))
    #mod_zey_only <- glmmTMB::glmmTMB(formula = form_zey_only, data = filtered_data, family = family(model))
    #aic_vals$zey_only <- AICc(mod_zey_only)
    
    #picta-only
    #form_picta_only <- as.formula(paste(response_var, "~", picta_predictor))
    #mod_picta_only <- glmmTMB::glmmTMB(formula = form_picta_only, data = filtered_data, family = family(model))
    #aic_vals$picta_only <- AICc(mod_picta_only)
    
    #lowest AIC
    lowest_model <- names(which.min(unlist(aic_vals)))
    aic_entry[[paste0("best_", thresh)]] <- lowest_model
  }
  
  aic_results[[species_name]] <- aic_entry
}

#convert to table
aic_table2 <- do.call(rbind, lapply(aic_results, as.data.frame))
rownames(aic_table2) <- NULL

#rename columns
colnames(aic_table2) <- c("Species", "lowest_all", "lowest_top50", "lowest_top25")
aic_table2

#write.csv(aic_table2, file = "mod_comp_table1.csv")
```

## Summary mammal table
```{r}
pruned_aggregated %>%
  group_by(Species, Family, TrophicLevel, ActivityCycle) %>%
  summarise(Total_Abundance = sum(Number.of.Animals, na.rm = TRUE), .groups = "drop") %>%
  mutate(
    Functional_Groups = paste(
      case_when(
        TrophicLevel == 1 ~ "Herbivore",
        TrophicLevel == 2 ~ "Omnivore",
        TrophicLevel == 3 ~ "Carnivore",
        TRUE ~ NA_character_
      ),
      case_when(
        ActivityCycle == 1 ~ "Nocturnal",
        ActivityCycle == 2 ~ "Crepuscular",
        ActivityCycle == 3 ~ "Diurnal",
        TRUE ~ NA_character_
      ),
      sep = ", "
    )
  ) %>%
  select(Species, Family, Functional_Groups, Total_Abundance) %>%
  arrange(desc(Total_Abundance)) %>%
  #write.csv(file = "MammalObsTable.csv", row.names = FALSE)
```

## Summary camera trap tables
```{r}
deployments <- read.csv("data/ofp_deployments-2021-11-04.csv", stringsAsFactors = FALSE)
partitions <- read.csv("data/CTlocations_partitions_May2024.csv", stringsAsFactors = FALSE)
by_day <- read.csv("data/30-daysamplingperiods-byday.csv", stringsAsFactors = FALSE)

#parse dates across df's 
parse_dates <- function(date_col) {
  parse_date_time(date_col, orders = c("mdy", "ymd"))
}

deployments <- deployments %>%
  mutate(Camera.Deployment.Begin.Date = parse_dates(Camera.Deployment.Begin.Date),
         Camera.Deployment.End.Date = parse_dates(Camera.Deployment.End.Date))

by_day <- by_day %>%
  mutate(date = parse_dates(date))

#expand deployments to daily scale 
deployments_daily <- deployments %>%
  filter(!is.na(Camera.Deployment.Begin.Date) & !is.na(Camera.Deployment.End.Date)) %>%
  mutate(date = map2(Camera.Deployment.Begin.Date, Camera.Deployment.End.Date, seq, by = "1 day")) %>%
  unnest(date) %>%
  select(locationID = Deployment.Location.ID, date) %>%
  distinct()

#link daily activity to sampling period and forest type
deployments_daily <- deployments_daily %>%
  inner_join(by_day, by = "date") %>%
  inner_join(partitions, by = "locationID")
```

Create camera trap by partition table
```{r}
#create grid of all combinations (64 periods x 12 partitions)
all_partitions <- partitions %>% 
  filter(habitat != "Kerangas") %>%
  select(habitat, partition) %>% 
  distinct()

all_periods <- unique(by_day$samplingperiod)

full_grid <- crossing(all_partitions, samplingperiod = all_periods)

#calculate active CT and CT nights per partition per sampling period
period_summary <- deployments_daily %>%
  group_by(habitat, partition, samplingperiod) %>%
  summarise(active_traps = n_distinct(locationID),
            trap_days = n(),
            .groups = "drop")

#complete grid to include 0s for inactive periods
period_summary_complete <- full_grid %>%
  left_join(period_summary, by = c("habitat", "partition", "samplingperiod")) %>%
  mutate(active_traps = replace_na(active_traps, 0),
         trap_days = replace_na(trap_days, 0))

#counts of total camera traps by partition
total_traps_count <- partitions %>%
  filter(habitat != "Kerangas") %>%
  group_by(habitat, partition) %>%
  summarise(N_camera_traps = n_distinct(locationID), .groups = "drop")

#camera trap nights over study period
total_days_count <- deployments_daily %>%
  filter(habitat != "Kerangas") %>%
  group_by(habitat, partition) %>%
  summarise(total_ct_days = n(), .groups = "drop")

#calculate mean and SD across sampling periods
metrics_across_periods <- period_summary_complete %>%
  group_by(habitat, partition) %>%
  summarise(mean_active = mean(active_traps),
            sd_active   = sd(active_traps),
            mean_days   = mean(trap_days),
            sd_days     = sd(trap_days),
            .groups     = "drop")

forest_order <- rev(c("Peat Swamp", "Freshwater Swamp", "Alluvial Bench", 
                  "Lowland Sandstone", "Lowland Granite", "Upland Granite", "Montane"))

#combine into final table
summary_table <- total_traps_count %>%
  left_join(metrics_across_periods, by = c("habitat", "partition")) %>%
  left_join(total_days_count, by = c("habitat", "partition")) %>%
  mutate(total_ct_days = replace_na(total_ct_days, 0)) %>%
  mutate(habitat = factor(habitat, levels = forest_order)) %>%
  arrange(habitat, partition) %>%
  transmute(
    `Forest Type` = habitat,
    `Forest Type Partition` = partition,
    `N Camera Traps` = N_camera_traps,
    `Active camera traps (Mean ± SD)` = paste0(round(mean_active, 2), " (±", round(sd_active, 2), ")"),
    `N camera Trap Nights` = total_ct_days,
    `Camera Trap Nights (Mean ± SD)` = paste0(round(mean_days, 2), " (±", round(sd_days, 2), ")")
  )

#write.csv(summary_table, "partition_summary_table.csv", row.names = FALSE)
```

check for continual coverage in camera traps by partition
```{r}
#shows which partitions during which sampling periods there was no CT surveying
period_summary_complete %>%
  filter(active_traps == 0) %>%
  select(habitat, partition, samplingperiod) %>%
  arrange(habitat, partition, samplingperiod) %>%
  print(n = Inf)
```

Gantt chart of CT deployments
```{r}
#prep data
gantt_data <- deployments %>%
  rename(locationID = Deployment.Location.ID) %>%
  inner_join(partitions, by = "locationID") %>%
  filter(!is.na(Camera.Deployment.Begin.Date) & !is.na(Camera.Deployment.End.Date)) %>%
  arrange(habitat, partition, Camera.Deployment.Begin.Date) %>%
  mutate(locationID = factor(locationID, levels = unique(locationID))) %>%
  filter(habitat != "Kerangas") %>%
  mutate(habitat = factor(habitat, levels = forest_order)) %>%
  arrange(habitat, partition, Camera.Deployment.Begin.Date)

#set custom color palette
forest_colors <- c(
  "Peat Swamp"        = "#401D00",
  "Freshwater Swamp"  = "#643E00",
  "Alluvial Bench"    = "#896007",
  "Lowland Sandstone" = "#AF8228",
  "Lowland Granite"   = "#D3A43B",
  "Upland Granite"    = "#F4C444",
  "Montane"           = "#FFDD2D")

ordered_colors <- forest_colors[forest_order]

#forcing labels to be horizontal and bold
num_facets <- length(forest_order)
horizontal_angles <- rep(0, num_facets)
bold_faces <- rep("bold", num_facets)

#plot
gantt <- ggplot(gantt_data) +
  geom_segment(aes(x = Camera.Deployment.Begin.Date, xend = Camera.Deployment.End.Date, 
                   y = locationID, yend = locationID, color = habitat), 
               size = 1.2, alpha = 0.8) +
  facet_grid2(habitat ~ ., scales = "free_y", space = "free_y",
              strip = strip_themed(
                text_y = elem_list_text(
                  color = ordered_colors,    
                  angle = horizontal_angles, 
                  face = bold_faces,         
                  size = 9))) +
  scale_x_date(date_breaks = "6 months", date_labels = "%b %Y") +
  scale_color_manual(values = forest_colors) +
  theme_minimal() +
  labs(x = "", y = "camera trap location ID") +
  theme(axis.text.y = element_text(size = 4), 
        axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
        strip.background = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        legend.position = "none")

#export 
#jpeg(filename = "gantt_chart.jpg",
 #    width = 6.5, height = 11.0,
  #   units = "in", res = 450, quality = 100)

#print(gantt)
#dev.off()
```
